---
title: "Untitled"
author: "Janine Quigley"
date: "2024-02-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(readxl)
library(readr)
library(mice)
library(ggplot2)
library(GGally)
library(arules)
library(ggfortify)
library(verification)
library(pROC)
library(pscl)
datafull = read_tsv("37138-0001-Data.tsv")

# USE ggsave()

```

#### Variable Breakdown:
 PTDTRACE, PESEX, PRTAGE, GEREG, PEMLR, GTCBSASZ, PEMEQ12, PEMEQ11, PESCHFT, PRCITSHP, PEMARITL, HEFAMINC, PRHRUSL,PEHRUSL1,PEEDUCA, PEC2Q1B, PEC1Q10A, PTC1Q10B, PEC2Q2B, PEMDQ4, PEMDQ6

```{r}
data = datafull %>% dplyr::select(PTDTRACE, PESEX, PRTAGE, GEREG, PEMLR, GTCBSASZ, 
  PEMEQ12, 
  PEMEQ11, 
  PESCHFT, 
  PRCITSHP, 
  PEMARITL, 
  HEFAMINC, 
  PRHRUSL,
  PEHRUSL1,
  PEEDUCA, 
  PEC2Q1B, 
  PEC1Q10A, 
  PTC1Q10B, 
  PEC2Q2B, 
  PEMDQ4, 
  PEMDQ6)

#-1 is blank/NA
```

```{r}

#filter based on dependent variable (museum visit Y/N)

data = data %>% filter(PEC1Q10A > 0) # leaves us with 8,711 obs


# new column names

col_names = c("Race", "Sex", "Age", "Region", "LaborForce", "CitySize", "ArtsInfo", "ArtsOpportunity", "Student", "CitizenshipStatus", "MaritalStatus", "FamilyIncome", "WrkHrsRange", "WeeklyHrsWrk", "Education", "VisitExhibit", "VisitMuseum", "NumMuseumVisit", "ExhibitVisitRange", "ArtsOrgSubscribe", "PurchaseArt")

colnames(data) = col_names
```

#### Deal with NA's

```{r}
data = subset(data, select=-c(VisitExhibit,ExhibitVisitRange))

cols_toNA = c("ArtsInfo", "ArtsOpportunity","Student","WrkHrsRange", "WeeklyHrsWrk", "ArtsOrgSubscribe", "PurchaseArt")
cols_tozero = c("NumMuseumVisit")

data = data %>% mutate_at(cols_toNA, ~ifelse(.< 0, NA,.))
data = data %>% mutate_at(cols_tozero, ~ifelse(.<0,0,.))

# Convert numeric columns to factors

factor_cols <- c("Race", "Sex", "Region", "FamilyIncome", "CitySize", "MaritalStatus", "Education", "CitizenshipStatus", "VisitMuseum", "ArtsOrgSubscribe", "PurchaseArt", "LaborForce", "ArtsInfo", "ArtsOpportunity", "Student", "WrkHrsRange")

data <- data %>%
  mutate_at(vars(factor_cols), factor)

subdata = data %>% dplyr::select(Race, Sex, Age, Region, LaborForce, CitySize, CitizenshipStatus, MaritalStatus, FamilyIncome, Education, VisitMuseum, NumMuseumVisit, ArtsOrgSubscribe, PurchaseArt)
data14 = na.omit(subdata) 
  # removed 5 variables with lots of NAs (Student, ArtsInfo, ArtsOpportunity, WrkHrsRange, WeeklyHrsWrk) - left with 830 x 14

subdata2 = data %>% dplyr::select(Race, Sex, Age, Region, LaborForce, CitySize, ArtsInfo, ArtsOpportunity, CitizenshipStatus, MaritalStatus, FamilyIncome, Education, VisitMuseum, NumMuseumVisit, ArtsOrgSubscribe, PurchaseArt) 
  # removed Student, WrkHrsRange, and WeeklyHrsWrk - left with 807 x 16
data16 = na.omit(subdata2)

subdata3 = data %>% dplyr::select(Race, Sex, Age, Region, LaborForce, CitySize, ArtsInfo, ArtsOpportunity, WrkHrsRange, WeeklyHrsWrk, CitizenshipStatus, MaritalStatus, FamilyIncome, Education, VisitMuseum, NumMuseumVisit, ArtsOrgSubscribe, PurchaseArt)
na.omit(subdata3) 
  #removed student - left with 468 x 18 ... is it enough?

```

##### Change Factor Names for Clarity

```{r}
racenames = c("White", "Black", "American Indian", "Asian", "Hawaiian/PI", "White/Black", "White/AI", "White/Asian", "White/H/PI", "Black/AI", "Asian/H/PI", "White/Black/AI", "White/Asian/H/PI")
levels(data14$Race) = racenames

levels(data14$Sex) = c("Male","Female")

regionnames = c("Northeast", "Midwest", "South", "West")
levels(data14$Region) = regionnames

laborforcenames = c("Employed(working)", "Employed(absent)", "Unemployed(layoff)", "Unemployed(looking)", "Retired", "Disabled", "Not In Labor Force")
levels(data14$LaborForce) = laborforcenames

citynames = c("Not Identified/Metropolitan", "100,000 - 249,999", "250,000 - 499,999", "500,000 - 999,999", "1,000,000 - 2,499,999", "2,500,000 - 4,999,999", "5,000,000+")
levels(data14$CitySize) = citynames

citizennames = c("Native - US Born", "Native - PR/other US Island Born", "Native - Born Abroad", "Foreign Born - US Citizen by Naturalization", "Foreign Born - Not US Citizen")
levels(data14$CitizenshipStatus) = citizennames

maritalnames = c("Married(spouse present)", "Married(spouse absent)", "Widowed", "Divorced", "Separated", "Never Married")
levels(data14$MaritalStatus) = maritalnames

incomenames = c("<$5,000", "$5,000 - 7,499", "$7,500 - 9,999", "$10,000 - 12,499", "$12,500 - 14,999", "$15,000 - 19,999", "$20,000 - 24,999", "$25,000 - 29,999", "$30,000 - 34,999", "$35,000 - 39,999", "$40,000 - 49,999", "$50,000 - 59,999", "$60,000 - 74,999", "$75,000 - 99,999", "$100,000 - 149,999", "$150,000+")
levels(data14$FamilyIncome) = incomenames

edunames = c("Less than 1st", "1st,2nd,3rd,or 4th", "5th/6th", "7th/8th", "9th", "10th", "11th", "12th(no diploma)", "Highschool/GED", "Some College", "Associate's(occupational)", "Associate's(academic)", "Bachelor's", "Master's", "Professional(MD/DDS/DVM)", "Doctorate(PhD/EdD)")
levels(data14$Education) = edunames

levels(data14$VisitMuseum) = c("Yes","No")
levels(data14$ArtsOrgSubscribe) = c("Yes","No")
levels(data14$PurchaseArt) = c("Yes","No")
```

###### multiple imputation

```{r}
imp = mice(data= data, m = 5, method = 'pmm', seed = 123)
summary(imp)

micedata = complete(imp) # second data, with all observations
```

### Summary Statistics

```{r}
write.csv(data14, "C:\\Users\\janin\\OneDrive\\Documents\\DA401\\MuseumResearchData.csv", row.names=FALSE)
# use 3,395 obs dataset
stats = summary(data14)
print(stats)


#stats3 = summary(micedata)
```

### Chi-Squared Tests

```{r}
chi1 = chisq.test(data14$Race, data14$VisitMuseum)
print(chi1) #  reject null hypo

chi2 = chisq.test(data14$Sex, data14$VisitMuseum)
print(chi2) #  reject null hypo

chi3 = chisq.test(data14$Region, data14$VisitMuseum)
print(chi3) # reject null hypo

chi4 = chisq.test(data14$LaborForce, data14$VisitMuseum)
print(chi4) # reject null hypo

chi5 = chisq.test(data14$CitySize, data14$VisitMuseum)
print(chi5) # reject null hypo

chi6 = chisq.test(data14$CitizenshipStatus, data14$VisitMuseum)
print(chi6) # CANNOT reject null hypo, and p > 0.05

chi7 = chisq.test(data14$MaritalStatus, data14$VisitMuseum)
print(chi7) # reject null hypo

chi8 = chisq.test(data14$FamilyIncome, data14$VisitMuseum)
print(chi8) # reject null hypo

chi9 = chisq.test(data14$Education, data14$VisitMuseum)
print(chi9) # reject null hypo

```

### Anova

```{r}
anova1 <- aov(NumMuseumVisit ~ Race + Sex +Age + Region + LaborForce + CitySize + CitizenshipStatus + MaritalStatus +  FamilyIncome +  Education, data=data14)
summary(anova1)
hist(data14$NumMuseumVisit)
```




### Linear Regression

```{r}
lm1 = lm(NumMuseumVisit ~ Race + Sex + Age + Region + LaborForce + CitySize + CitizenshipStatus + MaritalStatus + FamilyIncome + Education, data=data14)
summary(lm1)
residuals = resid(lm1)
# View the residuals
print(residuals)
autoplot(lm1)

plot(NumMuseumVisit ~ Race, data = data14)

ggplot(data14, aes(x = Race, y = NumMuseumVisit)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red") + labs(title="Race vs. Number of Museum Visits")
confint(lm1) #confidence interval

```

#### Logistic Regression

```{r}
log1 <- glm(VisitMuseum ~ Race + Sex + Age + Region + LaborForce + CitySize + CitizenshipStatus + MaritalStatus + FamilyIncome + Education, data = data14, family = "binomial")
summary(log1)
autoplot(log1)
predicted = predict(log1, type = "response")

# True labels
truelabels = data14$VisitMuseum

# Calculate ROC curve
roc_curve <- roc(truelabels, predicted)
plot.roc(roc_curve)
# Plot ROC curve
plot(roc_curve, main = "ROC Curve", col = "blue")

roc_curve <- roc(truelabels, predicted)

# Calculate AUC
auc_value <- auc(roc_curve)

```

##### correlation

```{r}
a = cbind(data14$Race, data14$Sex, data14$Age, data14$Region, data14$LaborForce, data14$CitySize, data14$CitizenshipStatus, data14$MaritalStatus, data14$FamilyIncome, data14$Education)
cormatrix = cor(a)
cormatrix
```


#### Poisson Model (to handle zero-heavy data)

```{r}
poisson1 = zeroinfl(NumMuseumVisit ~ Race + Sex + Age + Region + LaborForce + CitySize + CitizenshipStatus + MaritalStatus + FamilyIncome + Education | VisitMuseum, dist = 'poisson', data = data14)

with(data14, table(NumMuseumVisit,Race))
summary(poisson1)

negbi1 = zeroinfl(VisitMuseum ~ Race + Sex + Region + LaborForce + CitySize + CitizenshipStatus + MaritalStatus + FamilyIncome + Education | Race + Sex + Region + LaborForce + CitySize + CitizenshipStatus + MaritalStatus + FamilyIncome + Education, dist = 'negbin', data = data14)
summary(negbi1)
na.omit(data14)
```

#### decision trees, knn neighbors, random forest, 
